%\documentclass[12pt,twoside]{report}
\documentclass[12pt]{report}

%% last modification is June 4, 2016, Emma Pease

% note that the document can be single or double sided.  
% note that the Registrar's office now allows 10pt, 11pt, or 12pt

\usepackage{suthesis-2e}
% default is now online June/2016 version
%\usepackage[online]{suthesis-2e}
%\usepackage[hardcopy]{suthesis-2e}
% the following is for doing engineering theses. 
% I am definitely not sure of the wording on the signature page so check
%\usepackage[engineer]{suthesis-2e}

\usepackage[backend=bibtex,style=authoryear]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
\usepackage{amssymb, amsmath}
\usepackage{epsfig}
\usepackage{array}
\usepackage{ifthen}
\usepackage{color}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{chngcntr}
\usepackage{apptools}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

% one can change the default font to Times Roman but note that most
% ways of creating pdf files from latex automatically embed (which btw
% is a good idea even with the standard fonts)

    \title{Supervised Evaluation of Representations}
    \author{Charles Zheng}
    \dept{Statistics}
    \principaladviser{Trevor Hastie}
    \coprincipaladviser{Jonathan Taylor}
    \firstreader{Bradley Efron}
    \secondreader{Russell Poldrack (Psychology)}

%% one can also have a \thirdreader and \fourthreader

%% note that certain departments and types of theses have other requirements
%% For instance theses in the departments of 
%% Asian Languages
%% French and Italian
%% Spanish and Portuguese
%% need to define the \dept, \dualthesis, and the actual language
%\dualthesis
%\languagemajor{Chinese}
%% 
%% Those for Graduate Program in Humanities need to define 
%\humanitiesthesis
%\jointprogram{Arts and Crafts}
%% 
%% For submission to a committee or program (no department)
% \committeethesis
% \programthesis
%%
%% For School of Education or Business or Law
% \educationthesis
% \businessthesis
% \lawthesis  (law actually isn't listed in the official documents, 2013/1014)


\addbibresource{example.bib} % The filename of the bibliography





\newcommand{\tr}{\text{tr}}
\newcommand{\E}{\textbf{E}}
\newcommand{\diag}{\text{diag}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\Var}{\text{Var}}
\newcommand{\argmin}{\text{argmin}}
\newcommand{\Vol}{\text{Vol}}
\newcommand{\comm}[1]{}
\newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
\newcommand{\Cor}{\text{Cor}}
\newcommand{\bx}{\boldsymbol{x}}
\newcommand{\by}{\boldsymbol{y}}
\newcommand{\bX}{\boldsymbol{X}}
\newcommand{\bY}{\boldsymbol{Y}}
\newcommand{\bS}{\boldsymbol{S}}
\newcommand{\bu}{\boldsymbol{u}}
\newcommand{\bv}{\boldsymbol{v}}
\newcommand{\bU}{\boldsymbol{U}}
\newcommand{\bV}{\boldsymbol{V}}
\newcommand{\bB}{\boldsymbol{B}}
\newcommand{\bepsilon}{\boldsymbol{\epsilon}}
\newcommand{\bH}{\boldsymbol{H}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{definition}{Definition}[section]

\newcommand\crule[3][black]{\textcolor{#1}{\rule{#2}{#3}}}
\definecolor{color1}{RGB}{128,13,13}
\definecolor{color2}{RGB}{70,128,13}
\definecolor{color3}{RGB}{13,128,128}
\definecolor{color4}{RGB}{70,13,128}

\begin{document}

% for a variety of reasons this is an all in one document; however,
% when actually doing the thesis it is strongly recommended that each
% chapter be in a separate file and use \include to include in the
% main file.

%% the \beforepreface command produces the title page
%% in the online version it skips the copyright (page 2) and signature (page 3) pages 
%% in the non-online version these would be included
    \beforepreface


%% Abstract can be any number of pages
    \prefacesection{Abstract} Intuitively speaking, a good data
    representation reveals meaningful differences between inputs,
    while being relatively invariant to differences between inputs due
    to irrelevant factors or noise.  In this work, we consider criteria
    for formally defining the quality of a representation, which all
    make use of the availability of a \emph{response variable} $Y$ to
    distinguish between meaningful and meaningless variation.  Hence,
    these are criteria for \emph{supervised} evaluation of
    representations.  We consider three particular criteria: the mutual
    information between the representation and the reponse, the
    average accuracy of a randomized classification task, and the
    identification accuracy.  We discuss methods for estimating all
    three quantities, and also show how these three quantities are
    interrelated.  Besides the application of evaluating
    representations, our work also has relevance for estimation of
    mutual information in high-dimensional data, for obtaining
    performance guarantees for recognition systems, and for making
    statements about the generalizability of certain kinds of
    classification-based experiments which are found in neuroscience.

\newpage

%% one can also have a prefacesection that is a Preface instead of
%% Acknowledgements.   The thematic purpose is the same (thanks).
    \prefacesection{Acknowledgements} 

First of all, I would like to thank my co-advisors, Trevor Hastie and
Jonathan Taylor.  Trevor once told me that ``I am a bit of an unusual
student,'' but by the same token, Trevor is also an exceptional
advisor.  He really cares about making an impact in science and
industry, and also understands the difficulty of applied work in
statistics.  Many times when I was frustrated about the difficulty of
translating theoretical ideas into practically useful tools, Trevor
would inspire me to never give up on trying to bridge that gap between
theory and practice.  I also learned a great deal from him about how
to best present and communicate my ideas, which has helped immensely
in my collaborations.  I am no less fortunate to have had access to
Jonathan, whose wide-spanning technical expertise and passion for both
mathematical problem-solving and computation made him a valuable
resource to consult whenever I ran into any problem which seemed
beyond my grasp.  I can always count on Jonathan, with his sharp eye
for detail and his intensely curious mind, to provide numerous
corrections, suggestions, or connections anytime that I show him my
work in progress, and this applies especially to the thesis you now
hold in your hands (or--more likely--which is now displayed on your
monitor.)

In addition to Trevor and Jonathan, I had an additional, informal
advisor in my collaborator Yuval Benjamini.  Our extensive joint work,
which forms the basis for this thesis, was inspired by the class he
taught on statistical analysis of neuroscience data.  I must also
thank Yuval for his extraordinary support and assistance during my job
search.

I would also like to thank my committee members.  It has always been a
pleasure to work with Russell Poldrack, and I am especially grateful
to him for helping me find a job!  Bradley Efron has been an amazing
inspiration and support to me as a student, and I am truly honored to
have had the chance to discuss my research with one of the living
legends of statistics.  I am also deeply indebted to Tsachy Weissman
for being willing to lend his expertise in information theory, a field
ended up quite central to my research, despite my lack of knowledge in
the subject.  Of course, I would like to thank all of my committee
members for their valuable comments and questions they provided during
my oral defense, and I would like to thank my reading committee for
agreeing to read my thesis and for the feedback they have provided.

I have also relied extensively on many other faculty and staff members
at Stanford.  It has always been a pleasure to work with
Balasubramanian Narasimhan, or ``Naras,'' whose expertise at
statistical computing makes him a truly unique asset to the
department.  My time with Naras, both as his teaching assistant and as
his student, brought me to a higher level of expertise and
appreciation for the beauty of programming.  I am also grateful to Art
Owen, David Siegmund, Wing Wong, Robert Tibshirani, Persi Diaconis,
Richard Olshen, and Brian Wandell for many useful discussions.  I am
also greatly indebted to the department staff, not only for their work
in keeping the department running, but also for the help they've
personally given to me over the years for matters both great and
small: Cindy Kirby, Caroline Gates, Susie Ementon, Ellen van Stone,
Helen Tombropoulus, Heather Murthy, and Emily Lauderdale.

I would also like to thank the collaborators I have had over the
years: Ariel Rokem, Franco Pestilli, Sanmi Koyejo, Jingshu Wang, Arzav
Jain, Reza Zadeh, Bai Jiang, Nick Boyd, Rakesh Achanta, Qingyuan Zhao,
and Asaf Weinstein.  No less important are the friends at Stanford
whom I have to blame for making the five years of my Ph. D. seem to
pass in the blink of an eye: Rakesh Achanta, Chaojun Wang, Xin Zheng,
Christof Seiler, Asaf Weinstein, Veniamin Morgenstern, Pratheepa
Jeganathan, Lucy Xia, Kris Sankaran, Subharata Sen, Joey Arthur,
Junyang Qian, Nan Bi, Jun Yan, Yuming Kuang, Ya Le, Zeyu Zheng, Ziyu
Wang, Jiantao Jiao, Lucas Janson, Shuo Xie, Joshua Loftus, Murat
Erdogdu, Qianyuan Zhao, Paulo Orenstein, Keli Liu, Stephen Bates,
Hatef Monajemi, Edgar Dobriban, Cyrus DiCiccio, Oleg Lazarev, Zhou
Fan, Peter Diao, Fanya Becks, Kelvin Gu, Yan Chen, Xiaoyan Han, Haben
Michael, Jiada Liu, Frederik Mikelsen, Roman Croessmann, and many
others.

Finally, I am grateful to my family: my father, Buhong Zheng; my
mother, Xuemei Du; and my younger brother, Bruce Zheng, for supporting
me during my Ph. D. studies, and even volunteering as a test audience
for my job talk!
\newline

\noindent\textbf{Note on attribution}

The content in chapters 1,2, 4, and 5 is based on joint work with
Yuval Benjamini.  Chapter 3 is based on joint work with Yuval
Benjamini and Rakesh Achanta.  All theoretical results are due to the
author.

%% afterpreface produces a table of contents and any other tables
%% wanted. At the end pagenumbering changes from roman to arabic and
%% is restarted
    \afterpreface

\include{Chapters/Chapter1}
\include{Chapters/Chapter2} 
\include{Chapters/Chapter3}
\include{Chapters/Chapter4} 
\include{Chapters/Chapter5} 
\include{Chapters/Chapter6} 

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

\include{Appendices/AppendixA}
%\include{Appendices/AppendixB}
%\include{Appendices/AppendixC}
\include{Appendices/AppendixD}
%\include{Appendices/AppendixE}

 

\printbibliography[heading=bibintoc]

\end{document}

%% Normally the \chapter and any text would be in a separate file and included.

%% for Bibliography I would recommend a separate .bib file (bibtex) instead of inline as here.
%\nocite{astl2,perry1,pollardsag1}
%\nocite{*}
    \bibliographystyle{plain}
%    \bibliography{suthesis}

\begin{thebibliography}{1}

\bibitem{revelation}
{John of Patnos}.
\newblock Revelation of {St.} {John}.
\newblock In {\em Bible}. King James, 1611.

\bibitem{genesis}
Moses.
\newblock Genesis.
\newblock In {\em Bible}. King James, 1611.

\bibitem{preacher}
Preacher.
\newblock Ecclesiastes.
\newblock In {\em Bible}. King James, 1611.

\end{thebibliography}


\end{document}
